```{r setup}
#| include: false 
rm(list = ls())
library(tidyverse) 
library(ggfortify)
```

# Species

Allometric parameters and species functional traits for Paracou and Tapajos were used both as input to the TROLL assessment and as output to assess the functional composition of the forest. Many allometric parameters and functional traits did not exist and were inferred from dbh and height data or by using the relationship between traits with a predictive mean match.

## Allometric parameters

We took advantage of the TALLO database [@jucker2022] to infer species allometric parameters of the 496 species present in Amazonia (latitude between 10 and -18 and longitude between -78 and -39).

```{r tallodist}
#| message: false
#| fig-cap: "Available data in Amazonia."
ggplot(map_data("world"), aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(data = read_csv("data/species/tallo.csv") %>% 
               mutate(amazonia = ifelse(longitude <= -39 & 
                                          longitude >= -79 &
                                          latitude >= -18 &
                                          latitude <= 10, "amazonia", "other")),
  aes(x = longitude, y = latitude, col = amazonia)) +
  theme_bw() +
  geom_hline(yintercept = c(10, -18)) +
  geom_vline(xintercept = c(-39, -79)) +
  theme(axis.title = element_blank(), legend.position = "bottom") +
  scale_color_manual("", values = c("darkgreen", "black")) +
  coord_equal()
```

```{r hdbh}
#| message: false
#| fig-cap: "Height (m) and diameter (cm) relationship for observed amazonian trees."
read_csv("data/species/tallo.csv") %>% 
  filter(longitude <= -39, longitude >= -79, latitude >= -18, latitude <= 10) %>% 
  ggplot(aes(stem_diameter_cm, height_m)) +
  geom_point(aes(col = species), alpha = 0.25) +
  theme_bw() +
  geom_smooth() +
  scale_color_discrete(guide = "none") +
  xlab("diameter (cm)") + ylab("height (m)")
```

We inferred the following allometry using `brms`:

$$
\begin{array}{c} 
log(height) \sim N [ log(h_{max,s} \times \frac{dbh}{a_{h,s}+dbh^2}), \sigma^2] \\
h_{max,s} \sim N (h_{max}, \sigma_h^2) \\
a_{h,s} \sim N (a_{h}, \sigma_a^2)
\end{array}
$$

```{r fit}
#| eval: false
library(brms)
mdata <- read_csv("data/species/tallo.csv") %>% 
  filter(longitude <= -39, longitude >= -79, latitude >= -18, latitude <= 10) %>% 
  select(species, height_m, stem_diameter_cm) %>% 
  rename(h = height_m, d = stem_diameter_cm) %>% 
  mutate(d = d/100) %>% 
  mutate(species = ifelse(is.na(species), "unknown", species))
fit <- brm(bf(log(h) ~ log(hmax * (d/ (ah + d))), 
              hmax ~ 1 + 1|species,
              ah ~ 1 + 1|species,
              nl = TRUE), 
           prior = c(
             prior(normal(40, 10), lb = 10, nlpar = "hmax"),
             prior(normal(0.5, 0.5), lb = 0, nlpar = "ah")
             ),
           data = mdata, chains = 2, cores = 2, threads = 10)
save(fit, file = "outputs/brms_tallo.Rdata")
fit %>% 
  tidybayes::spread_draws(b_hmax_Intercept, r_species__hmax[species,],
               b_ah_Intercept, r_species__ah[species,]) %>% 
  mutate(ah = b_ah_Intercept + r_species__ah) %>% 
  mutate(hmax = b_hmax_Intercept + r_species__hmax) %>% 
  group_by(species) %>% 
  summarise(ah = median(ah), hmax = median(hmax)) %>% 
  filter(species != "unknown") %>% 
  mutate(species = gsub(".", " ", species, fixed = TRUE)) %>% 
  left_join(read_csv("data/species/tallo.csv") %>% 
  filter(longitude <= -39, longitude >= -79, latitude >= -18, latitude <= 10) %>% 
    select(family, genus, species) %>% unique()) %>% 
  write_tsv("outputs/tallo_pars.tsv")
```

We had good convergence and good post-predictive checks.

```{r rhat}
#| message: false
#| fig-cap: "Goodness of convergence assessed by Rhat. All Rhat are below 1.05 indicating a very good convergence of all chains."
load('outputs/brms_tallo.Rdata')
bayesplot::mcmc_rhat(bayesplot::rhat(fit$fit))
```

```{r ppc}
#| message: false
#| fig-cap: "Posterior post-predictive check of simulated response variable against observed values indicating a good representation of the whole distribution in posteriors."
load('outputs/brms_tallo.Rdata')
bayesplot::pp_check(fit)
```

We thus inferred 496 additional $a_h$, $h_{max}$ couple of parameters.

```{r tabfit}
load('outputs/brms_tallo.Rdata')
sjPlot::tab_model(fit)
```

```{r ahhmax}
#| fig-cap: "Inferred species allometric parameters hmax and ah and their relations."
fit %>% 
  tidybayes::spread_draws(b_hmax_Intercept, r_species__hmax[species,],
               b_ah_Intercept, r_species__ah[species,]) %>% 
  mutate(ah = b_ah_Intercept + r_species__ah) %>% 
  mutate(hmax = b_hmax_Intercept + r_species__hmax) %>% 
  group_by(species) %>% 
  summarise(ah = median(ah), hmax = median(hmax)) %>%
  filter(species != "unknown") %>% 
  ggplot(aes(ah, hmax)) +
  geom_point() +
  theme_bw() +
  xlab(expression(a[h])) + ylab(expression(h[max]))
```

```{r crown_radius_allometry, eval=F}
library(tidyverse)
library(sf)
library(brms)
x <- tibble(
  site = c("Paracou", "Tapajos"),
  latitude = c(5.27877, -2.85667),
  longitude = c(-52.92486, -54.9588900),
) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_buffer(10^6)
trees <- read_csv("data/species/tallo.csv") %>% 
  filter(longitude <= -39, longitude >= -79, latitude >= -18, latitude <= 10) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_intersection(x) %>% 
  filter(!is.na(crown_radius_m))
trees %>% 
  group_by(site) %>% 
  summarise(N = n())
ggplot(trees, aes(crown_radius_m, stem_diameter_cm, col = site)) +
  geom_point() +
  theme_bw() + 
  geom_smooth(method = "lm")
fit <- brm(bf(crown_radius_m ~ exp(cra + crb*log(stem_diameter_cm)), 
              cra ~ site,
              crb ~ site,
              nl = TRUE), 
           prior = c(
             prior(normal(0, 10), nlpar = "cra"),
             prior(normal(0, 10), nlpar = "crb")
             ),
           data = trees, chains = 2, cores = 2, threads = 10)
summary(fit)
```

## Functional traits

We further combined the obtained coefficient from `TALLO` with the following datasets:

-   @jucker2022
-   @maréchaux2015
-   @guillemot2022
-   @vleminckx2021
-   TROLL version 4 species data
-   Tapajos functional traits data retrieved by Jérôme Chave (pers. com.) from the TRY database [@Kattge2019]

TROLL version 4 species data were gathered from:

-   TROLL version 3 species (partly from @baraloto2010, described in @maréchaux2017, and available in \@schmitt2023 and )
-   @maréchaux2015
-   @maréchaux2019
-   @nemetschek2024
-   @schmitt2023a
-   Boisseaux *et al.* submitted
-   Ziegler *et al.* (which?)

All databases together included 20 traits across 2,921 species but with numerous missing data. Traits distribution were similar across datasets. However, for the same species, allometric parameters inferred from `TALLO` were lower than those inside TROLL V4 and leaf area from Guillemot was higher than those from TROLL and Vleminckx. As all Paracou TROLL species are in Vleminckx dataset and 109 species from Tapajos we will use only the species from Vleminckx to limit missing data in predictive mean matching (PMM) data imputation.

```{r spdata}
#| message: false
#| warning: false
# Maréchaux et al., (2015)
marechaux <- readxl::read_xls("data/species/marechaux.xls", "data") %>% 
  mutate(Pi_tlp = as.numeric(Pi_tlp)) %>% 
  separate(Binomial, c("genus", "species"), remove = FALSE) %>% 
  mutate(species = gsub("_", " ", Binomial)) %>% 
  rename(family = Family, TLP = Pi_tlp) %>% 
  dplyr::select(family, genus, species, TLP) %>% 
  mutate(dataset = "Maréchaux 2015") %>% 
  gather(trait, value, -dataset, -family, -genus, -species)
# Jucker et al., (2022) inferred
tallo <- read_tsv("outputs/tallo_pars.tsv") %>% 
  mutate(dataset = "Tallo inferred") %>% 
  gather(trait, value, -dataset, -family, -genus, -species)
# Vleminckx et al., (2021)
vleminckx <- readxl::read_xlsx("data/species/vleminckx.xlsx",
                  "App.S6-ok", skip = 3) %>% 
  mutate(species = sub("_", " ", Species)) %>% 
  mutate(species = gsub("[[:punct:]]", "", species)) %>% 
  rename(family = Family, genus = Genus, CC = Chlorophyll_content,
         LT = Thickness, LA = Leaf_Area, d13C = `13C`,
         BT = Trunk_bark_thickness) %>% 
    mutate(N = N/100, P = P/10^6, LMA = 1/SLA*1000) %>% 
  mutate(LMA = ifelse(LMA > 200, NA, LMA)) %>% 
  dplyr::select(family, genus, species, CC, LT, 
                Toughness, LA, LMA, C, N, d13C, Ca, P, K, 
                BT, WSG, SRL, SRTA) %>% 
   mutate(dataset = "Vleminckx 2021") %>% 
  gather(trait, value, -dataset, -family, -genus, -species)
# Guillemot et al., (2022)
guillemot <- readxl::read_xlsx("data/species/guillemot.xlsx", "Dataset") %>% 
  mutate(max_height = as.numeric(gsub(',', '.', max_height))) %>% 
  rename(LA = leaf_size, N = Leaf_N, P = Leaf_P, 
         WSG = Wood_density, hmax = max_height) %>% 
  mutate(TLP = ifelse(TLP < -3, NA, TLP)) %>% 
  mutate(P50 = ifelse(P50 < -10, NA, P50)) %>% 
  mutate(LA = ifelse(LA > 1000, NA, LA)) %>% 
  mutate(LMA = ifelse(LMA > 200, NA, LMA)) %>% 
  select(family, genus, species, TLP, P50, LMA, LA, WSG) %>% 
  mutate(dataset = "Guillemot 2022") %>% 
  gather(trait, value, -dataset, -family, -genus, -species)
# TROLL V4
troll <- rcontroll::TROLLv4_species %>% 
  rename_all(~ gsub("s_", "", .)) %>% 
  rename(species = name, N =Nmass, P = Pmass,
         WSG = wsg, TLP = tlp, LA = leafarea) %>% 
   mutate(species = gsub("_", " ", species)) %>% 
  select(-seedmass, -regionalfreq) %>% 
  left_join(select(vleminckx, family, genus, species) %>% unique()) %>% 
  mutate(dataset = "TROLL V4") %>% 
  gather(trait, value, -dataset, -family, -genus, -species)
# Tapajos
tapajos <- vroom::vroom("data/species/tapajos_try.csv") %>% 
  separate(species, c("genus", "Species"), remove = FALSE) %>% 
  mutate(family = BIOMASS::getTaxonomy(genus, findOrder = FALSE)$family)  %>% 
  filter(!is.na(Species)) %>% 
  filter(!is.na(family)) %>% 
  mutate(LMA = 1/SLA*1000, N = `N_try(mg/g)`/1000, WSG = wood_density,
         P = `P_try(mg/g)`/1000, dbhmax = dbh_max/100) %>% 
  select(family, genus, species, LMA, N, P, WSG, dbhmax) %>% 
  mutate(dataset = "Tapajos TRY") %>% 
  gather(trait, value, -dataset, -family, -genus, -species)
# All
raw <- bind_rows(marechaux, 
                 tallo,
                 vleminckx, 
                 guillemot, 
                 troll,
                 tapajos) %>% 
  na.omit()  
filtered <- raw %>% 
  filter(species %in% vleminckx$species) %>% 
  group_by(family, genus, species, dataset, trait) %>% 
  summarise(value = median(value)) %>% 
  group_by(family, genus, species, trait) %>% 
  summarise(value = median(value)) %>% 
  pivot_wider(names_from = trait, values_from = value) %>% 
  ungroup()
```

```{r traittab}
#| message: false
filtered %>% 
  select(-family, -genus) %>% 
  gather(trait, value, -species) %>% 
  na.omit() %>% 
  group_by(trait) %>% 
  summarise(N = n()) %>% 
  knitr::kable(caption = "Available functional traits data per sepcies.")
```

```{r vlem}
#| message: false
options(knitr.kable.NA = '')
raw %>% 
  group_by(species, dataset) %>% 
  summarise(value = 1) %>% 
  pivot_wider(names_from = dataset, values_from = value) %>% 
  mutate(`Vleminckx 2021` = ifelse(is.na(`Vleminckx 2021`), "Vleminckx: no", "Vleminckx: yes")) %>% 
  ungroup() %>% 
  select(-species) %>% 
  gather(dataset, value, -`Vleminckx 2021`) %>% 
  na.omit() %>% 
  group_by(`Vleminckx 2021`, dataset) %>% 
  summarise(N = n()) %>% 
  pivot_wider(names_from = `Vleminckx 2021`, values_from = N) %>% 
  knitr::kable(caption = "Available functional traits data per species depending on the dataset and its overlap with Vleminchkx's data.")
```

```{r traitdistraw}
#| fig-cap: "Raw distributions of functional traits data per dataset origin."
raw %>% 
  mutate(trait = gsub("_", " ", trait)) %>% 
  ggplot(aes(dataset, value, col = dataset)) +
  geom_boxplot() +
  facet_wrap(~ trait, scales = "free") +
  theme_bw() +
  theme(axis.title = element_blank(), axis.text.x = element_blank(),
        legend.position = "bottom") 
```

```{r traitdistrawcomp}
#| message: false
#| fig-cap: "Species traits means distributions per datasets in function of their median across datasets."
raw %>% 
  select(species, dataset, trait, value) %>% 
  group_by(species, dataset, trait) %>% 
  summarise(value = median(value)) %>% 
  group_by(species, trait) %>%
  filter(n() > 1) %>%
  mutate(median = median(value, na.rm = TRUE)) %>% 
  ggplot(aes(median, value, col = dataset)) +
  geom_point() +
  facet_wrap(~ trait, scales = "free") +
  theme_bw() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r rawcorr}
#| fig-cap: "Species traits pairwise correlations."
filtered %>% 
  select(-family, -genus, -species) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  corrplot::corrplot(type = "upper", diag = FALSE)
```

## Imputation

Missing data were too important to use random forest for imputation with `missForest` even with phylogenetic eigen values. Thus, we used simple predictive mean matching (PMM) data imputation with `mice` without phylogenetic eigen values. However, phylogenetic eigen values could also been used with `phylomice` (not done here). As we focused on species in common with Vlemincks data we only had to impute ah, dbhmax, hmax and TLP with values from Maréchaux, Tapajos and inferred with `TALLO`. Obtained imputed distributions matched well the observed distribution of the raw data and imputed functional space match the one of the raw data (PCA). The PCA also revealed similar functional spaces between species from Paracou and Tapajos, whereas Tapajos functional space is a bit wider.

```{r mice}
#| eval: false
imputation <- filtered %>% 
  mice::mice(maxit = 100)
complete(imputation) %>% 
  gather(trait, value, -family, -genus, -species) %>% 
  mutate(type = "imputed") %>% 
  bind_rows(gather(filtered, trait, value, -family, -genus, -species) %>% 
              mutate(type = "raw")) %>% 
  na.omit() %>% 
  write_tsv("outputs/imputation.tsv")
```

```{r dens}
#| message: false
#| fig-cap: "Traits density distribution before and after species trait values imputation by predictive means matching."
t <- vroom::vroom("outputs/imputation.tsv") %>% 
  filter(trait %in% c("ah", "dbhmax", "hmax", "TLP")) %>% 
  mutate(trait_long = recode(trait, "ah" = "a[h]", "dbhmax" = "dbh[max]", "hmax" = "h[lim]", "TLP" = "psi[~tlp]"))
labels <- t %>% 
  group_by(trait, type) %>% 
  summarise(N = n()) %>% 
  pivot_wider(names_from = type, values_from = N) %>% 
  mutate(label = paste0("Raw: ", raw, "\nImputed:", imputed)) %>% 
  ungroup() %>% 
  mutate(value = c(-2, 0.1, 1, 30),
         density = c(0.25, 5, 1, 0.05)) %>% 
  mutate(trait_long = recode(trait, "ah" = "a[h]", "dbhmax" = "dbh[max]", "hmax" = "h[lim]", "TLP" = "psi[~tlp]"))
g <- ggplot(t, aes(value)) +
  geom_density(aes(col = type)) +
  geom_text(data = labels, aes(label = label, y = density)) +
  theme_bw() +
  facet_wrap(~ trait_long, scales = "free", labeller = label_parsed) +
  scale_color_discrete("") +
  theme(legend.position = "bottom", axis.title = element_blank())
ggsave("figures/fs01.png", g, dpi = 300, width = 5, height = 5, bg = "white")
g
```

```{r pca1}
#| message: false
#| fig-cap: "Traits distribution in the first plan of a PCA before and after species trait values imputation by predictive means matching."
data <- vroom::vroom("outputs/imputation.tsv") %>% 
  filter(trait %in% troll$trait) %>% 
  pivot_wider(names_from = trait, values_from = value) %>% 
  na.omit()
autoplot(princomp(select(data, -family, -genus, -species, -type), cor = TRUE), 
         data = data, alpha = 0.25, col = "type",
         loadings.label.size = 6,
         loadings.label.colour = 'red', loadings.label.vjust = 1.1,
         loadings.label.repel	= TRUE,
         loadings = T, loadings.label = T, loadings.colour = 'red') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme_bw()
```

```{r pca2}
#| fig-cap: "Traits distribution in the first plan of a PCA among sites after imputation by predictive means matching."
data2 <- data %>% 
  filter(type == "imputed") %>% 
  mutate(tapajos = (species %in% tapajos$species)) %>% 
  mutate(paracou = (species %in% troll$species)) %>% 
  filter(tapajos | paracou) %>% 
  mutate(forest = ifelse(tapajos, "tapajos", "paracou")) %>% 
  mutate(forest = ifelse(paracou & tapajos, "both", forest)) %>% 
  select(-tapajos, -paracou, -type)
autoplot(princomp(select(data2, -family, -genus, -species, -forest), cor = TRUE), 
         data = data2, alpha = 0.5, col = "forest",
         loadings.label.size = 6,
         loadings.label.colour = 'red', loadings.label.vjust = 1.1,
         loadings.label.repel	= TRUE,
         loadings = T, loadings.label = T, loadings.colour = 'red') +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'black', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'black', linetype = "dotted") +
  theme_bw() +
  stat_ellipse(aes(col = forest))
```

```{r data2troll}
#| eval: false
data2 %>% 
  filter(forest != "paracou") %>% 
  select(-forest, -family, -genus) %>% 
  mutate(seedmass = 1, regionalfreq = 1/n()) %>% 
  mutate(species = gsub(" ", "_", species)) %>% 
  rename(s_name = species, s_LMA = LMA, s_Nmass = N, s_Pmass = P, s_wsg = WSG,
         s_dbhmax = dbhmax, s_hmax = hmax, s_ah = ah, 
         s_seedmass = seedmass, s_regionalfreq = regionalfreq, 
         s_tlp = TLP, s_leafaera = LA) %>% 
  write_tsv("outputs/tapajos_species_troll.tsv")
data2 %>% 
  filter(forest != "tapajos") %>% 
  select(-forest, -family, -genus) %>% 
  mutate(seedmass = 1, regionalfreq = 1/n()) %>% 
  mutate(species = gsub(" ", "_", species)) %>% 
  rename(s_name = species, s_LMA = LMA, s_Nmass = N, s_Pmass = P, s_wsg = WSG,
         s_dbhmax = dbhmax, s_hmax = hmax, s_ah = ah, 
         s_seedmass = seedmass, s_regionalfreq = regionalfreq, 
         s_tlp = TLP, s_leafaera = LA) %>% 
  write_tsv("outputs/paracou_species_troll.tsv")
```

## Functional composition

Forest functional composition was simply assessed using functional trait distribution in each plots (repeted lines in Paracou data) at the species level in Paracou and genus level in Tapajos.

```{r dist}
#| eval: false
traits <- bind_rows(read_tsv("outputs/tapajos_species_troll.tsv"),
          read_tsv("outputs/paracou_species_troll.tsv")) %>% 
  unique() %>% 
  rename_all(~ gsub("s_", "", .)) %>% 
  select(-seedmass, -regionalfreq) %>% 
  rename(species = name) %>% 
  mutate(species = gsub("_", " ", species)) %>% 
  gather(trait, trait_value, -species)
paracou <- read_tsv("outputs/inventories.tsv") %>% 
  filter(site == "Paracou", type == "default") %>% 
  mutate(plot = ifelse(site == "Tapajos", 1, plot)) %>% 
  filter(type == "default") %>% 
  select(site, plot, species) %>% 
  left_join(traits) %>% 
  na.omit()
tapajos <- read_tsv("outputs/inventories.tsv") %>% 
  filter(site == "Tapajos", type == "default") %>% 
  mutate(site = "Tapajos", plot = 1) %>% 
  separate(species, "species") %>% 
  select(plot, site, species) %>% 
      left_join(traits %>% 
  separate(species, "species") %>% 
  group_by(species, trait) %>% 
  summarise_all(mean)) %>% 
  na.omit()
bind_rows(paracou, tapajos) %>% 
  write_tsv("outputs/functional_composition.tsv")
```

```{r paracoudist}
#| message: false
#| fig-cap: "Functional composition at the Paracou site expressed in terms of density distribution per trait. The analyses have been done at the species level in Paracou."
read_tsv("outputs/functional_composition.tsv") %>% 
  filter(site == "Paracou") %>% 
  ggplot(aes(trait_value, group = plot)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ trait, scales = "free") +
  theme_bw() +
  ggtitle("Paracou", "546,498 inds represented on 577,761 (94.59%)") +
  theme(axis.title = element_blank(), legend.position = "bottom") +
  scale_color_discrete("")
```

```{r tapajosdist}
#| message: false
#| fig-cap: "Functional composition at the Tapajos site expressed in terms of density distribution per trait. The analyses have been done at the genus level in Tapajos."
read_tsv("outputs/functional_composition.tsv") %>% 
  filter(site == "Tapajos") %>% 
  ggplot(aes(trait_value, group = plot)) +
  geom_density() +
  facet_wrap(~ trait, scales = "free") +
  theme_bw() +
  ggtitle("Tapajos Genus level", "19,188 inds represented on 19,499 (98.65%)") +
  theme(axis.title = element_blank(), legend.position = "bottom") +
  scale_color_discrete("")
```
