```{r setup}
#| include: false 
rm(list = ls())
library(tidyverse) 
```

# Climate

Climatic data for Paracou and Tapajos used for the TROLL evaluation.

## ERA5-Land

ERA5-Land is a global reanalysis dataset [@muñoz-sabater2021]. ERA5-Land at Paracou was retrieved from the Climate Data Store (CDS) using the R package `rcontroll` [@schmitt2023]. ERA5-Land at Tapajos was retrieved from [Open-Meteo](https://open-meteo.com/) using mamba environment and python script available in the data folder. ERA5-Land was mainly used to validate FLUXNET 2015 and gap-filling.

## FLUXNET 2015

FLUXNET 2015 [@PastorelloGilberto2020] at Paracou and Tapajos was retrieved freely from [FLUXNET](https://fluxnet.org/data/fluxnet2015-dataset/) (the manifest is available in data). They are mainly used for TROLL model forcing but also include fluxes for TROLL simulations validation. However, FLUXNET 2015 precipitation data at Tapajos are not reliable, we thus replaced it with precipitation from ERA5-Land.

> The FLUXNET2015 Dataset includes data collected at sites from multiple [regional flux networks](https://fluxnet.org/about/regional-networks/). The preparation of this FLUXNET Dataset has been possible thanks only to the efforts of many scientists and technicians around the world and the coordination among teams from regional networks.

PI: Damien Bonal (Paracou), Mike Goulden (Tapajos), Scott Saleska (Tapajos)

```{r fluxnet}
#| message: false
#| eval: false
fluxnet <- list(
  "Paracou" = "data/climate/FLX_GF-Guy_FLUXNET2015_SUBSET_HH_2004-2014_2-4.csv",
  "Tapajos" = "data/climate/FLX_BR-Sa1_FLUXNET2015_SUBSET_HR_2002-2011_1-4.csv"
) %>% 
  lapply(vroom::vroom, na = "-9999") %>% 
  bind_rows(.id = "site") %>% 
  select(site, TIMESTAMP_START, TIMESTAMP_END, P_F, SW_IN_F, 
         TA_F, VPD_F, WS_F, GPP_NT_VUT_REF, LE_F_MDS) %>% 
  mutate(time = as_datetime(as.character(as.POSIXlt(as.character(TIMESTAMP_START), 
                                                    format = "%Y%m%d%H%M")))) %>% 
  rename(rainfall = P_F, snet = SW_IN_F, temperature = TA_F, vpd = VPD_F, ws = WS_F) %>% 
  mutate(vpd = vpd/10) %>% 
  select(site, time, rainfall, snet, temperature, vpd, ws)
fluxnet_tapajos <- filter(fluxnet, site == "Tapajos")
era <- vroom::vroom("data/climate/era5_tapajos.csv") %>% 
  filter(year(date) > 2001) %>% 
  mutate(time = as_datetime(date), rainfall = precipitation) %>% 
  select(time, rainfall)
fluxnet_tapajos <- fluxnet_tapajos %>% 
  select(-rainfall) %>% 
  left_join(era)
time_freq <- 0.5 # TROLL v4
start <- as_datetime(as_date(min(fluxnet_tapajos$time))) # to start from 00:30
stop <- max(fluxnet_tapajos$time) + 30*60 # to stop at 23:00
time <- seq(start, stop, by = 60 * 60 * time_freq)
data <- left_join(tibble(time = time),
    fluxnet_tapajos,
    by = "time"
  ) %>%
  group_by(date = as_date(time)) %>%
  mutate(across(c(temperature, vpd, ws),
                ~ zoo::na.spline(., time, na.rm = FALSE))) %>% 
  mutate(across(c(snet),
                ~ zoo::na.approx(., time, na.rm = FALSE))) %>% 
  ungroup() %>% 
  select(-date) %>% 
  mutate(across(c(snet), ~ ifelse(is.na(.), 0, .))) %>% 
  mutate(across(c(snet), ~ ifelse(. < 0, 0, .))) %>% 
  mutate(site = "Tapajos")
bind_rows(filter(fluxnet, site == "Paracou"), data) %>% 
  write_tsv("outputs/fluxnet_climate.tsv")
```

## Comparisons

ERA5-Land and FLUXNET2015 showed high correlation which comforted us in using ERA5-Land to gap-fill FLUXNET2015 precipitaiton data at Tapajos.

```{r comp}
#| message: false
#| fig-cap: "Comparisons of monthly means over ten years for Paracou between FLUXNET2015 and ERA5-Land weather data."
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, fluxnet, -time, -site) %>% 
  left_join(read_tsv("data/climate/era5_Paracou.tsv") %>% 
              gather(variable, era, -time) %>% 
              mutate(site = "Paracou")) %>% 
  na.omit() %>% 
  group_by(site, variable, date = floor_date(time, "month")) %>% 
  select(-time) %>% 
  mutate_all(~ ifelse(variable == "rainfall", sum(.), .)) %>% 
  summarise_all(mean) %>% 
  ggplot(aes(fluxnet, era)) +
  geom_point(alpha = 0.5, col = "lightgrey") +
  facet_wrap(~ variable, scales = "free") +
  geom_abline(linetype = "dashed") +
  ggpubr::stat_cor() +
  theme_bw() +
  ggtitle("Monthly variations")
```

```{r tap_rain}
#| message: false
#| warning: false
#| eval: false
#| include: false
# https://noaa-ghcn-pds.s3.amazonaws.com/index.html#csv/by_station/
g1 <- read_tsv("outputs/fluxnet_climate.tsv") %>% 
  filter(site == "Tapajos") %>% 
  select(time, rainfall) %>% 
  group_by(date = date(time)) %>% 
  summarise(rainfall = sum(rainfall, na.rm = TRUE)) %>% 
  left_join(
    read_csv("data/fluxes/km67.eddyflux.seasonal.txt") %>% 
      mutate(date = as_date(paste0(YYYY, "-", MM, "-", DD))) %>% 
      select(date, rainBelterra, rainTRMM) %>% 
      rename(Belterra = rainBelterra, TRMM = rainTRMM)
  ) %>% 
  ggplot(aes(rainfall, Belterra)) +
  geom_point() +
  geom_abline() +
  theme_bw()  +
  ggpubr::stat_cor() +
  xlab("Precipitation from ERA5-Land [mm/day]") +
  ylab("Precipitation from Belterra [mm/day]") +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Daily ERA5-Land vs. Belterra", "446 days")
g2 <- read_tsv("outputs/fluxnet_climate.tsv") %>% 
  filter(site == "Tapajos") %>% 
  select(time, rainfall) %>% 
  group_by(date = date(time)) %>% 
  summarise(rainfall = sum(rainfall, na.rm = TRUE)) %>% 
  group_by(date = floor_date(date, "15 days")) %>% 
  summarise(rainfall = sum(rainfall)) %>% 
  left_join(
    read_csv("data/fluxes/km67.eddyflux.seasonal.txt") %>% 
      mutate(date = as_date(paste0(YYYY, "-", MM, "-", DD))) %>% 
      select(date, rainBelterra, rainTRMM) %>% 
      rename(Belterra = rainBelterra, TRMM = rainTRMM) %>% 
      group_by(date = floor_date(date, "15 days")) %>% 
      summarise(TRMM = sum(TRMM))
  ) %>% 
  ggplot(aes(rainfall, TRMM)) +
  geom_point() +
  geom_abline() +
  theme_bw()  +
  ggpubr::stat_cor() +
  xlab("Precipitation from ERA5-Land [mm/16-day]") +
  ylab("Precipitation from TRMM [mm/16-day]") +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("16-day ERA5-Land vs. TRMM", "460 16-day")
g3 <- read_csv("data/climate/tapajos_chirps.csv") %>% 
      rename(date = time, rainfall = pr) %>% 
  select(date, rainfall) %>% 
  filter(rainfall > 0.1) %>% 
  left_join(
    read_csv("data/fluxes/km67.eddyflux.seasonal.txt") %>% 
      mutate(date = as_date(paste0(YYYY, "-", MM, "-", DD))) %>% 
      select(date, rainBelterra, rainTRMM) %>% 
      rename(Belterra = rainBelterra, TRMM = rainTRMM)
  ) %>% 
  ggplot(aes(rainfall, Belterra)) +
  geom_point() +
  geom_abline() +
  theme_bw()  +
  ggpubr::stat_cor() +
  xlab("Precipitation from CHIRPS [mm/day]") +
  ylab("Precipitation from Belterra [mm/day]") +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("Daily CHIRPS vs. Belterra", "446 days")
g4 <- read_csv("data/climate/tapajos_chirps.csv") %>% 
  rename(date = time, rainfall = pr) %>% 
  select(date, rainfall) %>% 
  filter(rainfall > 0.1) %>% 
  group_by(date = floor_date(date, "15 days")) %>% 
  summarise(rainfall = sum(rainfall)) %>% 
  left_join(
    read_csv("data/fluxes/km67.eddyflux.seasonal.txt") %>% 
      mutate(date = as_date(paste0(YYYY, "-", MM, "-", DD))) %>% 
      select(date, rainBelterra, rainTRMM) %>% 
      rename(Belterra = rainBelterra, TRMM = rainTRMM) %>% 
      group_by(date = floor_date(date, "15 days")) %>% 
      summarise(TRMM = sum(TRMM))
  ) %>% 
  ggplot(aes(rainfall, TRMM)) +
  geom_point() +
  geom_abline() +
  theme_bw()  +
  ggpubr::stat_cor() +
  xlab("Precipitation from CHIRPS [mm/16-day]") +
  ylab("Precipitation from TRMM [mm/16-day]") +
  scale_x_log10() +
  scale_y_log10() +
  ggtitle("16-day CHIRPS vs. TRMM", "460 16-day")
g <- cowplot::plot_grid(g1, g2, g3, g4)
ggsave("test.png", g)
```

```{r par_rain}
#| message: false
#| warning: false
#| fig-cap: "Paracou precipitation across products."
pr_paracou <- read_tsv("outputs/fluxnet_climate.tsv") %>% 
  filter(site == "Paracou") %>% 
  select(time, rainfall) %>% 
  group_by(date = date(time)) %>% 
  summarise(pr_fluxnet = sum(rainfall)) %>% 
  left_join(read_tsv("data/climate/era5_Paracou.tsv") %>% 
              select(time, rainfall) %>% 
              group_by(date = date(time)) %>% 
              summarise(pr_era5 = sum(rainfall, na.rm = TRUE))) %>% 
  left_join(read_csv("data/climate/paracou_chirps.csv") %>% 
              group_by(date = date(time)) %>% 
              summarise(pr_chirps = sum(pr, na.rm = TRUE)))
dat_daily <- pr_paracou %>% 
  rename("ERA5-Land" = pr_era5, CHIRPS = pr_chirps) %>% 
  gather(source, pr, -date, -pr_fluxnet) %>% 
  mutate(type = "day")
dat_pentad <- pr_paracou %>% 
  group_by(date = floor_date(date, "5 days")) %>% 
  summarise_all(sum) %>% 
  rename("ERA5-Land" = pr_era5, CHIRPS = pr_chirps) %>% 
  gather(source, pr, -date, -pr_fluxnet) %>%
  mutate(type = "5-days")
dat_biweek <- pr_paracou %>% 
  group_by(date = floor_date(date, "15 days")) %>% 
  summarise_all(sum) %>% 
  rename("ERA5-Land" = pr_era5, CHIRPS = pr_chirps) %>% 
  gather(source, pr, -date, -pr_fluxnet) %>%
  mutate(type = "15-days")
dat_month <- pr_paracou %>% 
  group_by(date = floor_date(date, "month")) %>% 
  summarise_all(sum) %>% 
  rename("ERA5-Land" = pr_era5, CHIRPS = pr_chirps) %>% 
  gather(source, pr, -date, -pr_fluxnet) %>% 
  mutate(type = "month")
bind_rows(dat_daily, dat_pentad, dat_biweek, dat_month) %>% 
  mutate(type = factor(type, levels = c("day", "5-days", "15-days", "month"))) %>% 
  mutate(pr = ifelse(pr < 1, 0, pr)) %>% 
  mutate(pr_fluxnet = ifelse(pr_fluxnet < 1, 0, pr_fluxnet)) %>% 
  ggplot(aes(pr_fluxnet, pr, col = source)) +
  geom_abline() +
  geom_point(alpha = 0.2) +
  theme_bw() +
  coord_equal() +
  scale_y_log10() +
  scale_x_log10() +
  geom_smooth(method = "lm", se = FALSE) +
  ggpubr::stat_cor(aes(label = ..rr.label..)) +
  facet_wrap(~ type) +
  xlab("Precipitation FLUXNET2015") +
  ylab("Precipitation other products") +
  scale_color_discrete("")
```

## Input data

TROLL input data for climate includes:

-   daily rainfall (mm)

-   half-hourly day net radiation (snet, W/m2)

-   half-hourly day temperature (°C)

-   half-hourly day vapor pressure deficit (vpd, Pa)

-   half-hourly day wind speed (m/s)

We thus need to defined day limits in time at a half-hourly precision. For that we used daily variations of snet. We thus used 7-19h for Paracou and 6-18h for Tapajos.

```{r snetdailyvar}
#| message: false
#| fig-cap: "Half-hourly variaations of day net radiation across ten years at Paracou and Tapajos to define day limits to be used in TROLL simulations."
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  group_by(site, time = hms::as_hms(time)) %>% 
    summarise(l = quantile(snet, 0.025, na.rm = TRUE), 
            m = mean(snet, na.rm = TRUE), 
            h = quantile(snet, 0.975, na.rm = TRUE)) %>% 
  mutate(time = as.numeric(time)/(60*60)) %>% 
  ggplot(aes(time, m, col = site, fill = site)) +
  geom_vline(data = data.frame(time = c(7, 19, 6, 18), 
                               site = c("Paracou", "Paracou",
                                        "Tapajos", "Tapajos")), 
             aes(xintercept = time, col = site), 
             linetype = "dashed", linewidth = 1.5) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  geom_point() +
  theme_bw() +
  theme(axis.title = element_blank(), legend.position = "bottom") +
  ggtitle("Half-hourly variations") +
  scale_fill_discrete("") +
  scale_color_discrete("") +
  geom_vline(xintercept = c(7, 19), linetype = "dotted")
```

## Dry seasons

For representation and discussion purposes we defined dry season as 15-days period below 50 mm in means across years (corresponding to month below 100 mm from @bonal2008 but see @fu2013 for a more precise approach). We thus defined a 4-month dry season in Paracou from 8/1 to 12/1 and an almost 5-month dry season in Tapajos from 6/15 to 11/1.

```{r ds}
#| message: false
#| fig-cap: "Fortnightly means of precipitation in Paracou and Tapajos across ten years used to define dry site's climatological dry season."
read_tsv("outputs/fluxnet_climate.tsv") %>%
  group_by(site, date = floor_date(time, "15 days")) %>% 
  summarise(rainfall = sum(rainfall, na.rm = TRUE)) %>% 
  mutate(week = week(date), year = year(date)) %>% 
  mutate() %>% 
  ggplot(aes(week,  rainfall)) +
  geom_rect(aes(xmin = week, xmax = end), fill  = "#fff4e0",
           ymin = -Inf, ymax = Inf, alpha = 0.5,
           data = data.frame(week = week(as_date(c("2000-8-1", "2000-06-15"))),
                             rainfall = -Inf,
                             end = week(as_date(c("2000-12-1", "2000-11-1"))),
                             site = c("Paracou", "Tapajos"))) +
  geom_line(aes(group = year), col = "grey") +
  theme_bw() +
  facet_wrap(~ site) +
  geom_smooth(col = "black", se = FALSE) +
  scale_x_continuous(breaks = week(as_date(paste("2000-", 1:12, "-01"))),
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  theme(axis.title = element_blank()) +
  geom_hline(yintercept = 50, linetype = "dashed")
```

## Variations in time

For data exploration purposes, we represented climate variations at multiple time steps (year, month, half-month, week, day, and half-hour), but ultimately TROLL is forced with half-hourly variations except for precipitation which is daily.

```{r yearly}
#| message: false
#| warning: false
#| fig-cap: "Yearly means of climate variables at Paracou and Tapajos across ten years."
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, value, -site, -time) %>% 
  group_by(site, variable, date = date(time)) %>% 
  select(-time) %>% 
   mutate(value = ifelse(variable %in% c("rainfall", "snet"),
                           sum(value, na.rm = T), value)) %>% 
  summarise_all(mean) %>% 
  group_by(site, variable, year = year(date)) %>% 
  select(-date) %>% 
  mutate(value = ifelse(variable %in% c("rainfall"),
                           sum(value, na.rm = T), value)) %>% 
    summarise(l = quantile(value, 0.025, na.rm = TRUE), 
            m = mean(value, na.rm = TRUE), 
            h = quantile(value, 0.975, na.rm = TRUE)) %>% 
  ggplot(aes(year, m, col = site, fill = site)) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  geom_point() +
  theme_bw() +
  theme(axis.title = element_blank(), legend.position = c(0.8, 0.2)) +
  ggtitle("Yearly variations of daily means") +
  scale_fill_discrete("") +
  scale_color_discrete("") +
  facet_wrap(~ variable, scales = "free_y")
```

```{r monthly}
#| message: false
#| fig-cap: "Mean annual cycle of monthly means of climate variables at Paracou and Tapajos across ten years"
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, value, -site, -time) %>% 
  group_by(site, variable, date = floor_date(time, "month")) %>% 
  select(-time) %>% 
  mutate(value = ifelse(variable %in% c("rainfall"),
                           sum(value, na.rm = T), value)) %>% 
  summarise_all(mean) %>%
  group_by(month = month(date)) %>% 
  mutate(week = mean(week(date))) %>%  
  group_by(site, variable, week) %>% 
  summarise(l = quantile(value, 0.025, na.rm = TRUE), 
            m = mean(value, na.rm = TRUE), 
            h = quantile(value, 0.975, na.rm = TRUE)) %>% 
  ggplot(aes(week, m, col = site, fill = site)) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  geom_point() +
  theme_bw() +
  theme(axis.title = element_blank(), legend.position = c(0.8, 0.2)) +
  ggtitle("Monthly means variations") +
  scale_fill_discrete("") +
  scale_color_discrete("") +
  facet_wrap(~ variable, scales = "free_y") +
    scale_x_continuous(breaks = week(as_date(paste("2000-", 1:12, "-01"))),
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"))
```

```{r 15days}
#| message: false
#| fig-cap: "Mean annual cycle of fortnightly means of climate variables at Paracou and Tapajos across ten years."
floor_biweek <- function(full) 
  as_date(paste0(year(full), "-", month(full), "-", ifelse(day(full) > 15, 15, 1)))
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, value, -site, -time) %>% 
  group_by(site, variable, date = floor_date(time, "15 days")) %>% 
  select(-time) %>% 
  mutate(value = ifelse(variable %in% c("rainfall"),
                           sum(value, na.rm = T), value)) %>% 
  summarise_all(mean) %>%
  mutate(week = week(floor_biweek(date))) %>%  
  group_by(site, variable, week) %>% 
  summarise(l = quantile(value, 0.025, na.rm = TRUE), 
            m = mean(value, na.rm = TRUE), 
            h = quantile(value, 0.975, na.rm = TRUE)) %>% 
  ggplot(aes(week, m, col = site, fill = site)) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_smooth(se = FALSE) +
  theme_bw() +
  theme(axis.title = element_blank(), legend.position = c(0.8, 0.2)) +
  ggtitle("15-days means variations") +
  scale_fill_discrete("") +
  scale_color_discrete("") +
  facet_wrap(~ variable, scales = "free_y") +
    scale_x_continuous(breaks = week(as_date(paste("2000-", 1:12, "-01"))),
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"))
```

```{r week}
#| message: false
#| fig-cap: "Mean annual cycle of weekly means of climate variables at Paracou and Tapajos across ten years."
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, value, -site, -time) %>% 
  group_by(site, variable, date = floor_date(time, "week")) %>% 
  select(-time) %>% 
  mutate(value = ifelse(variable %in% c("rainfall"),
                           sum(value, na.rm = T), value)) %>% 
  summarise_all(mean) %>%
  group_by(site, variable, week = week(date)) %>% 
  mutate(value = ifelse(variable %in% c("rainfall"),
                           sum(value, na.rm = T), value)) %>% 
  summarise(l = quantile(value, 0.025, na.rm = TRUE), 
            m = mean(value, na.rm = TRUE), 
            h = quantile(value, 0.975, na.rm = TRUE)) %>% 
  ggplot(aes(week, m, col = site, fill = site)) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  geom_point() +
  theme_bw() +
  theme(axis.title = element_blank(), legend.position = c(0.8, 0.2)) +
  ggtitle("Weekly means variations") +
  scale_fill_discrete("") +
  scale_color_discrete("") +
  facet_wrap(~ variable, scales = "free_y") +
    scale_x_continuous(breaks = week(as_date(paste("2000-", 1:12, "-01"))),
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"))
```

```{r day}
#| message: false
#| fig-cap: "Mean annual cycle of daily means of climate variables at Paracou and Tapajos across ten years."
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, value, -site, -time) %>% 
  group_by(site, variable, date = floor_date(time, "day")) %>% 
  select(-time) %>% 
  mutate(value = ifelse(variable %in% c("rainfall"),
                           sum(value, na.rm = T), value)) %>% 
  summarise_all(mean) %>%
  group_by(site, variable, day = yday(date)) %>% 
  summarise(l = quantile(value, 0.025, na.rm = TRUE), 
            m = mean(value, na.rm = TRUE), 
            h = quantile(value, 0.975, na.rm = TRUE)) %>% 
  ggplot(aes(day, m, col = site, fill = site)) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank(), legend.position = c(0.8, 0.2)) +
  ggtitle("Weekly means variations") +
  scale_fill_discrete("") +
  scale_color_discrete("") +
  facet_wrap(~ variable, scales = "free_y") +
    scale_x_continuous(breaks = yday(as_date(paste("2000-", 1:12, "-01"))),
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"))
```

```{r halfhourly}
#| message: false
#| fig-cap: "Mean daily cycle of half-hourly means of climate variables at Paracou and Tapajos across ten years."
read_tsv("outputs/fluxnet_climate.tsv") %>% 
  gather(variable, value, -site, -time) %>% 
  filter(variable %in% c("snet", "temperature", "vpd", "ws")) %>% 
  group_by(site, variable, time = hms::as_hms(time)) %>% 
    summarise(l = quantile(value, 0.025, na.rm = TRUE), 
            m = mean(value, na.rm = TRUE), 
            h = quantile(value, 0.975, na.rm = TRUE)) %>% 
  mutate(time = as.numeric(time)/(60*60)) %>% 
  ggplot(aes(time, m, col = site, fill = site)) +
  geom_vline(data = data.frame(time = c(7, 19, 6, 18), 
                               site = c("Paracou", "Paracou",
                                        "Tapajos", "Tapajos")), 
             aes(xintercept = time, col = site), linetype = "dashed") +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  geom_point() +
  theme_bw() +
  facet_wrap(~ variable, scales = "free") +
  theme(axis.title = element_blank(), legend.position = "bottom") +
  ggtitle("Half-hourly variations") +
  scale_fill_discrete("") +
  scale_color_discrete("")
```
