```{r setup}
#| include: false 
library(tidyverse) 
library(knitr)
```

# Structure v4

Calibration of forest structure parameters (crown radius allometry and mortality) in TROLL 4.0.1 (as mortality differs between TROLL 3.1.8 and 4.0.1):

-   $m$ : minimal death rate in death per year calibrated with TROLL 3.1.8 to 0.055 both at Paracou and Tapajos
-   $a_{CR}$ : crown radius intercept calibrated with TROLL 3.1.8 to 1.85 at Paracou and to 2.50 at Tapajos
-   $b_{CR}$ : crown radius slope calibrated with TROLL 3.1.8 to 0.4445 at Paracou and to 0.8150 at Tapajos

## Parameters spaces

We prepared the parameter space for calibration defining low limit, high limit, and step for a, m, and the error of b for Paracou and Tapajos independently using calibration of forest crown radius allometry from TROLL 3.1.8 as prior and tests with TROLL 4.0.1 for the mortality prior. We obtained a calibration grid of **810 simulations**. Dashed lines shows prior values. The last figure shows the corresponding projected allometries of crown radius.

```{r prepparamspace}
pars_par <- data.frame(site = "Paracou",
                       parameter = c("a", "error_b", "m"),
                       low = c(1.6, -6*0.05, 0.03),
                       high = c(2.0, -2*0.05, 0.05),
                       by = c(0.05, 0.05, 0.0025))
pars_tap <- data.frame(site = "Tapajos",
                       parameter = c("a", "error_b", "m"),
                       low = c(2.3, -6*0.05, 0.03),
                       high = c(2.7, -2*0.05, 0.05),
                       by = c(0.05, 0.05, 0.0025))
pars <- bind_rows(pars_par, pars_tap)
grid_par <- data.frame(a = seq(pars_par$low[1], pars_par$high[1], by = pars_par$by[1])) %>% 
  mutate(error_b = list(seq(pars_par$low[2], pars_par$high[2], by = pars_par$by[2]))) %>% 
  unnest(error_b) %>% 
  mutate(b = -0.39 + 0.57*a + error_b) %>% 
  mutate(m = list(seq(pars_par$low[3], pars_par$high[3], by = pars_par$by[3]))) %>% 
  unnest(m) %>% 
  mutate(site = "Paracou")
grid_tap <- data.frame(a = seq(pars_tap$low[1], pars_tap$high[1], by = pars_tap$by[1])) %>% 
  mutate(error_b = list(seq(pars_tap$low[2], pars_tap$high[2], by = pars_tap$by[2]))) %>% 
  unnest(error_b) %>% 
  mutate(b = -0.39 + 0.57*a + error_b) %>% 
  mutate(m = list(seq(pars_tap$low[3], pars_tap$high[3], by = pars_tap$by[3]))) %>% 
  unnest(m) %>% 
  mutate(site = "Tapajos")
grid <- bind_rows(grid_par, grid_tap)
```

```{r parstab}
kable(pars)
```

```{r abplot}
grid %>% 
  ggplot(aes(a, b, col = site)) +
  geom_point() +
  geom_vline(aes(xintercept = a, col = site), linetype = "dashed", 
             data = data.frame(a = c(1.85, 2.50), site = c("Paracou", "Tapajos"))) +
  geom_hline(aes(yintercept = b, col = site), linetype = "dashed", 
             data = data.frame(b = c(0.4445, 0.8150), site = c("Paracou", "Tapajos"))) +
  geom_abline(intercept = -0.37, slope = 0.57, col = "black") +
  theme_bw() +
  xlab(expression(a[CR])) +
  ylab(expression(b[CR]))
```

```{r amplot}
grid %>% 
  ggplot(aes(a, m, col = site)) +
  geom_point() +
  geom_vline(aes(xintercept = a, col = site), linetype = "dashed", 
             data = data.frame(a = c(1.85, 2.50), site = c("Paracou", "Tapajos"))) +
  geom_hline(yintercept = 0.040, linetype = "dashed") +
  theme_bw() +
  xlab(expression(a[CR])) +
  ylab(expression(m))
```

```{r crallomplot}
grid %>% 
  mutate(dbh = list(seq(1, 150, 1)/100)) %>% 
  unnest(dbh) %>% 
  mutate(cr = exp(a+b*log(dbh))) %>% 
  ggplot(aes(dbh, cr)) +
  geom_line(aes(col = a, group = paste(a, b))) +
  theme_bw() +
  scale_colour_viridis_c(expression(a[CR])) +
  xlab("Diameter (m)") + ylab("Crown radius (m)") +
  geom_line(col = "black", linetype = "dashed",
            data = bind_rows(
              data.frame(site = "Paracou", 
                         dbh = seq(1, 150, 1)/100) %>% 
              mutate(cr = exp(1.85+0.4445*log(dbh))),
              data.frame(site = "Tapajos", 
                         dbh = seq(1, 150, 1)/100) %>% 
              mutate(cr = exp(2.50+0.8150*log(dbh)))
            )
              
            ) +
  facet_wrap(~ site)
```

## Observations

We used different evaluation metrics per site to rely on a single data source and minimise discrepancies.

For Tapajos and according to @rice2004 :

-   470 live trees per hectare with diameter at breast height (dbh) ≥10
-   Aboveground live biomass was 143.7 ± 5.4 Mg C/ha
-   Stem density -0.058 (±0.003) dbh_cm + 2.912 (±0.076) for dbh \< 40
-   Stem density -0.016 (±0.001) dbh_cm + 1.451 (±0.04) for dbh \> 40
-   Stem density is shown below

For Paracou according to the inventories:

-   612 ± 37 live trees per hectare with diameter at breast height (dbh) ≥10 (from inventory)
-   Basal area was 30.8 ± 0.836 m2/ha
-   Stem density is shown below

```{r obs}
#| message: false
abd_par <- read_tsv("outputs/inventories.tsv", show_col_types = FALSE) %>% 
  filter(site == "Paracou", type == "default") %>% 
  filter(dbh > 10) %>% 
  group_by(plot) %>% 
  summarise(abundance = n()/unique(plot_size)) %>% 
  ungroup()
# paste(round(median(abd_par$abundance)), "±", round(sd(abd_par$abundance)))
# read_tsv("outputs/inventories.tsv") %>% 
#   filter(site == "Paracou", type == "default", dbh >= 10) %>% 
#   group_by(plot) %>% 
#   summarise(ba = sum((dbh/2)^2*pi)/10^4/unique(plot_size)) %>% 
#   summarise(ba_mean = median(ba), ba_sd = sd(ba))
obs_tap <- data.frame(dbh_class = seq(10, 150, 5)) %>% 
  mutate(abundance = ifelse(dbh_class < 40, 
                            10^(-0.058*dbh_class+2.912), 
                            10^(-0.016*dbh_class+1.451))) %>% 
  mutate(site = "Tapajos")

obs_par <- read_tsv("outputs/inventories.tsv", show_col_types = FALSE) %>% 
  filter(site == "Paracou", type == "default") %>% 
  filter(dbh > 10) %>% 
  mutate(dbh_class = cut(dbh, 
                         breaks = c(seq(10, 155, by = 5)), 
                         labels = c(seq(10, 150, by = 5)))) %>% 
  mutate(dbh_class = as.numeric(as.character(dbh_class))) %>% 
    group_by(plot, dbh_class) %>% 
  summarise(abundance = n()/unique(plot_size)) %>% 
  group_by(dbh_class) %>% 
  summarise(abundance = median(abundance)) %>% 
  mutate(site = "Paracou")
# obs <- bind_rows(obs_tap, obs_par)
obs <- bind_rows(obs_tap, obs_par)
ggplot(obs, (aes(dbh_class, abundance, col = site))) +
  geom_line() +
  theme_bw() +
  xlab("DBH [ cm ]") + ylab(expression(Abundance~"["~ha^{-~1}~"]")) +
  ggtitle(paste("Total:", round(sum(obs$abundance))))
```

## Parameters effect

```{r calib}
#| eval: false
bind_rows(
  read_tsv("results/calib_str4_Paracou.tsv"),
  read_tsv("results/calib_str4_Tapajos.tsv")
) %>% 
  mutate(dbh_class = cut(dbh_cm, 
                         breaks = seq(10, 205, by = 5), 
                         labels = seq(10, 200, by = 5),
                         include.lowest = TRUE)) %>% 
  mutate(dbh_class = as.numeric(as.character(dbh_class))) %>% 
  group_by(site, a, b, m, dbh_class) %>% 
  summarise(abundance = n()/4, 
            agb = sum(agb)/4, 
            ba = sum((dbh_cm/2)^2*pi)/10^4/4) %>% 
  write_tsv("outputs/calib_str4.tsv")
```

### $a_{CR}$

```{r aCR}
#| message: false
#| warning: false
dat <-  read_tsv("outputs/calib_str4.tsv") %>% 
  group_by(site, a, b, m) %>% 
  summarise(agb = sum(agb)/10^3/2,
            ba = sum(ba),
            abundance = sum(abundance)) %>% 
  filter(m == 0.040)
cowplot::plot_grid(
 dat %>% 
  ggplot(aes(a, b, col = agb)) +
  geom_point() +
  facet_wrap(~ site, scales = "free") +
  theme_bw() +
  scale_color_viridis_c("AGB") +
  xlab(expression(a[CR|m==0.04])) + ylab(expression(b[CR|m==0.04])),
  dat %>% 
  ggplot(aes(a, b, col = ba)) +
  geom_point() +
  facet_wrap(~ site, scales = "free") +
  theme_bw() +
  scale_color_viridis_c("BA") +
  xlab(expression(a[CR|m==0.04])) + ylab(expression(b[CR|m==0.04])),
   dat %>% 
  ggplot(aes(a, b, col = abundance)) +
  geom_point() +
  facet_wrap(~ site, scales = "free") +
  theme_bw() +
  scale_color_viridis_c("Abundance") +
  xlab(expression(a[CR|m==0.04])) + ylab(expression(b[CR|m==0.04])), 
 nrow = 3)
```

```{r aCR2}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  group_by(site, a, b, m) %>% 
  summarise(abundance = sum(abundance), 
            agb = sum(agb)/10^3/2) %>% 
  ggplot(aes(abundance, agb, col = a)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(expression(a[CR])) +
  facet_wrap(~ site)
```

```{r aCRdist}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  filter(m == 0.040) %>% 
  mutate(berr = -0.39 + 0.57*a - b) %>% 
  filter(berr > 0.1, berr < 0.3) %>% 
  ggplot(aes(dbh_class, abundance)) +
  geom_line(aes(col = a, group = paste(a, b, m))) +
  geom_point(data = filter(obs, dbh_class < 150), col = "red") +
  scale_y_log10(labels = scales::comma) +
  xlab("DBH [ cm ]") + ylab(expression(Abundance~"["~ha^{-~1}~"]")) +
  scale_color_viridis_c(expression(a[CR|m==0.04|berr==0.02])) +
  theme_bw() +
  facet_wrap(~ site, nrow = 2)
```

### $b_{CR}$

```{r bCR}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  group_by(site, a, b, m) %>% 
  summarise(abundance = sum(abundance), 
            agb = sum(agb)/10^3/2) %>% 
  ggplot(aes(abundance, agb, col = b)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(expression(b[CR])) +
  facet_wrap(~ site)
```

### $m$

```{r m2}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  group_by(site, a, b, m) %>% 
  summarise(abundance = sum(abundance), 
            agb = sum(agb)/10^3/2) %>% 
  ggplot(aes(abundance, agb, col = m)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(expression(m)) +
  facet_wrap(~ site)
```

## Best parameters

### Paracou

```{r best_par}
#| message: false
#| warning: false
eval_par <- read_tsv("outputs/calib_str4.tsv") %>% 
  filter(site == "Paracou") %>% 
  select(site, a, b, m) %>% 
  unique() %>% 
  mutate(dbh_class = list(seq(10, 150, by = 5))) %>% 
  unnest(dbh_class) %>% 
  left_join(read_tsv("outputs/calib_str4.tsv")) %>% 
  select(-agb) %>% 
  mutate_at(c("abundance", "ba"), ~ifelse(is.na(.), 0, .)) %>% 
  rename(simulated = abundance) %>% 
  left_join(rename(obs, observed = abundance)) %>% 
  mutate(observed = ifelse(is.na(observed), 0, observed)) %>% 
  group_by(site, a, b, m) %>% 
  summarise(rmsep_dist = sqrt(mean((observed - simulated)^2)),
            rmsep_ba = sqrt((30.8 - sum(ba))^2),
            rmsep_abund = sqrt((612 - sum(simulated))^2)
            ) %>% 
  ungroup() %>% 
  mutate(all_rmsep_norm = rmsep_dist/30.5 + rmsep_ba/30.8 + rmsep_abund/612) %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m))
best_par <- eval_par %>% 
  group_by(site) %>% 
  slice_min(all_rmsep_norm, n = 5)
```

```{r besttab_par}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m)) %>% 
  left_join(best_par) %>% 
  filter(!is.na(all_rmsep_norm)) %>% 
  group_by(site, a, b, m) %>% 
  summarise(abundance = sum(abundance), ba = sum(ba)) %>% 
  gather(variable, value, -site) %>% 
  group_by(site, variable) %>% 
  summarise(median = median(value), minimum = min(value), maximum = max(value)) %>% 
  knitr::kable(caption = "Best structure parameters according to calibration with inventory data.")
```

```{r bestfig_par}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m)) %>% 
  left_join(best_par %>% slice_min(all_rmsep_norm)) %>% 
  filter(!is.na(all_rmsep_norm)) %>% 
  ggplot(aes(dbh_class, abundance)) +
  geom_line(aes(group = sim)) +
  geom_point(data = obs %>% filter(site == "Paracou"), col = "red") +
  scale_y_log10(labels = scales::comma) +
  xlab("DBH [ cm ]") + ylab(expression(Abundance~"["~ha^{-~1}~"]")) +
  theme_bw() +
  facet_wrap(~ site)
```

### Tapajos

```{r best_tap}
#| message: false
#| warning: false
eval_tap <- read_tsv("outputs/calib_str4.tsv") %>% 
  filter(site == "Tapajos") %>% 
  select(site, a, b, m) %>% 
  unique() %>% 
  mutate(dbh_class = list(seq(10, 150, by = 5))) %>% 
  unnest(dbh_class) %>% 
  left_join(read_tsv("outputs/calib_str4.tsv")) %>% 
  select(-ba) %>% 
  mutate_at(c("abundance", "agb"), ~ifelse(is.na(.), 0, .)) %>% 
  rename(simulated = abundance) %>% 
  left_join(rename(obs, observed = abundance)) %>% 
  mutate(observed = ifelse(is.na(observed), 0, observed)) %>% 
  group_by(site, a, b, m) %>% 
  summarise(rmsep_dist = sqrt(mean((observed - simulated)^2)),
            rmsep_agb = sqrt((143.7 - sum(agb)/10^3/2)^2),
            rmsep_abund = sqrt((470 - sum(simulated))^2)
            ) %>% 
  ungroup() %>% 
  mutate(all_rmsep_norm = rmsep_dist/16.2 + rmsep_agb/143.7 + rmsep_abund/470) %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m))
best_tap <- eval_tap %>% 
  group_by(site) %>% 
  slice_min(all_rmsep_norm, n = 5)
```

```{r besttab_tap}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m)) %>% 
  left_join(best_tap) %>% 
  filter(!is.na(all_rmsep_norm)) %>% 
  group_by(site, a, b, m) %>% 
  summarise(abundance = sum(abundance), agb = sum(agb)/10^3/2) %>% 
  gather(variable, value, -site) %>% 
  group_by(site, variable) %>% 
  summarise(median = median(value), minimum = min(value), maximum = max(value)) %>% 
  knitr::kable(caption = "Best structure parameters according to calibration with inventory data.")
```

```{r bestfig_tap}
#| message: false
#| warning: false
read_tsv("outputs/calib_str4.tsv") %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m)) %>% 
  left_join(best_tap %>% slice_min(all_rmsep_norm)) %>% 
  filter(!is.na(all_rmsep_norm)) %>% 
  ggplot(aes(dbh_class, abundance)) +
  geom_line(aes(group = sim)) +
  geom_point(data = obs %>% filter(site == "Tapajos"), col = "red") +
  scale_y_log10(labels = scales::comma) +
  xlab("DBH [ cm ]") + ylab(expression(Abundance~"["~ha^{-~1}~"]")) +
  theme_bw() +
  facet_wrap(~ site)
```

### All

```{r bestab2}
bind_rows(
  best_par %>% slice_min(all_rmsep_norm),
  best_tap %>% slice_min(all_rmsep_norm)
) %>% 
  select(site, a, b, m) %>% 
  mutate(berr = b + 0.39 - 0.57*a) %>% 
  kable()
```

```{r bestfig2}
#| message: false
#| warning: false
g <- read_tsv("outputs/calib_str4.tsv") %>% 
  mutate(sim = paste0(site, "_", a, "_", b, "_", m)) %>% 
  left_join(
    bind_rows(
      best_tap %>% slice_min(all_rmsep_norm),
      best_par %>% slice_min(all_rmsep_norm)
    )
  ) %>% 
  filter(!is.na(all_rmsep_norm)) %>% 
  ggplot(aes(dbh_class, abundance)) +
  geom_line(aes(group = sim, col = "TROLL")) +
  geom_point(data = obs, aes(col = "field")) +
  xlab("DBH [ cm ]") + ylab(expression(Abundance~"["~ha^{-~1}~"]")) +
  theme_bw() +
  facet_wrap(~ site) +
  scale_colour_manual("", values = c("#3f3f3f", "#0da1f8")) +
  theme(legend.position = "bottom") +
  geom_text(aes(col = "field", label = lab), 
            data = data.frame(dbh_class = 80, abundance = 200,
                              lab = c("470 tress/ha & 143.7 MgC/ha",
                                      "612 tress/ha & 30.8 m2/ha"),
                              site = c("Tapajos", "Paracou"))) +
  geom_text(aes(col = "TROLL", label = lab), 
            data = data.frame(dbh_class = 80, abundance = 180,
                              lab = c("471 tress/ha & 144.9 MgC/ha",
                                      "622 tress/ha & 30.5 m2/ha"),
                              site = c("Tapajos", "Paracou")))
ggsave("figures/f01.png", g, dpi = 300, width = 8, height = 5, bg = "white")
g
```
