```{r setup}
#| include: false 
library(tidyverse) 
cols <- c("sim" = "#0da1f8", "obs" = "#3f3f3f")
```

# Phenology

Calibration of the new leaf phenology modules against observed litterfall in Paracou and Tapajos. Parameters include:

-   a: the leaf resistance parameter
-   b: the tree height susceptibility parameter
-   delta: the base variation in residence time multiplying factor for drought shedding

```{r calib}
#| eval: false
prep_sim <- function(f) {
  t <- vroom::vroom(f)  %>% 
    select(iter, litterfall) %>% 
    mutate(file = f) %>% 
    mutate(file = gsub("results/calib_pheno/", "", file)) %>% 
    mutate(file = gsub("_0_sumstats.txt", "", file)) %>% 
    separate(file, c("site", "a0", "b0", "delta"), "_", convert = T) %>% 
    mutate(date = as_date(ifelse(site == "Tapajos", 
                                 "2002/01/01", "2004/01/01")) + iter) %>% 
    mutate(litterfall = litterfall*365) %>% 
    select(-iter) %>% 
    left_join(vroom::vroom("outputs/litter.tsv") %>% 
                rename(observed = litterfall))
  t$obs_group <- NA
  t$obs_group[!is.na(t$dt)] <- 1:length(t$dt[!is.na(t$dt)])
  t %>% 
    fill(obs_group, .direction = "up") %>% 
    mutate(dt_full = dt) %>% 
    fill(dt_full, .direction = "up") %>% 
    group_by(obs_group) %>% 
    filter(n() >= dt_full) %>% 
    mutate(simulated = ifelse(!is.na(dt), 
                              zoo::rollmeanr(litterfall, unique(dt_full), fill = NA), 
                              NA)) %>% 
    ungroup() %>% 
    select(site, a0, b0, delta, date, dt, observed, simulated) %>% 
    na.omit()
}

files <- list.files("results/calib_pheno", full.names = T, pattern = "sumstats.txt")
names(files) <- list.files("results/calib_pheno", full.names = F, pattern = "sumstats.txt")
raw <- files %>% 
  lapply(prep_sim)  %>% 
  bind_rows()
write_tsv(raw, "outputs/calib.tsv")
```

```{r calib_test_ll_isa}
#| eval: false
prep_sim <- function(f) {
  t <- vroom::vroom(f)  %>% 
    select(iter, litterfall) %>% 
    mutate(iter = iter - 219000 + 365*10 + 1) %>% 
    filter(iter > 0) %>% 
    mutate(file = f) %>% 
    mutate(file = gsub("results/test_ll_isa/", "", file)) %>% 
    mutate(file = gsub("_testLL_0_sumstats.txt", "", file)) %>% 
    mutate(file = gsub("TROLLv4_", "", file)) %>% 
    rename(site = file) %>% 
    mutate(date = as_date(ifelse(site == "Tapajos", 
                                 "2002/01/01", "2004/01/01")) + iter) %>% 
    mutate(litterfall = litterfall*365) %>% 
    select(-iter) %>% 
    left_join(vroom::vroom("outputs/litter.tsv") %>% 
                rename(observed = litterfall))
  t$obs_group <- NA
  t$obs_group[!is.na(t$dt)] <- 1:length(t$dt[!is.na(t$dt)])
  t %>% 
    fill(obs_group, .direction = "up") %>% 
    mutate(dt_full = dt) %>% 
    fill(dt_full, .direction = "up") %>% 
    group_by(obs_group) %>% 
    filter(n() >= dt_full) %>% 
    mutate(simulated = ifelse(!is.na(dt), 
                              zoo::rollmeanr(litterfall, unique(dt_full), fill = NA), 
                              NA)) %>% 
    ungroup() %>% 
    select(site, date, dt, observed, simulated) %>% 
    na.omit()
}

files <- list.files("results/test_ll_isa", full.names = T, pattern = "sumstats.txt")
names(files) <- list.files("results/test_ll_isa", full.names = F, pattern = "sumstats.txt")
raw <- files %>% 
  lapply(prep_sim)  %>% 
  bind_rows()
write_tsv(raw, "outputs/calib_test_ll_isa.tsv")
```

```{r indices}
#| message: false
sim_metrics <- read_tsv("outputs/calib.tsv") %>% 
  gather(type, litterfall, -site, -a0, -b0, -delta, -date, -dt) %>% 
  group_by(site, a0, b0, delta, year = year(date), type) %>% 
    summarise(peak = mean(c(litterfall[which.max(litterfall)],
                          litterfall[which.max(litterfall)-1],
                          litterfall[which.max(litterfall)+1]), na.rm = TRUE),
            peak_date = date[which.max(litterfall)],
            basal = mean(as.numeric(month(date) %in% 1:4)*litterfall, na.rm = TRUE)) %>%  
  mutate(peak_day = yday(peak_date), ratio = peak / basal) %>% 
  filter(basal > 0, peak_day > 100)
```

## Parameters effect

The leaf resistance parameter which is limiting the potential at turgor loss point (TLP, see methods) is manly limiting the height of litterfall peak in dry season with a peak up to 30 MgC/ha/yr with a at 0.01 against no peak with high values above 0.3. An higher leaf resistance above 0.1 also results in a latter peak. Globally, the less resistant are the leaves of the trees the higher is the litterfall in dry conditions.

```{r a0}
#| message: false
#| warning: false
#| fig-cap: "Effect of the leaf resistance parameter a on the annual cycle from fortnightly litterfall for Paracou and Tapajos."
ga1 <- read_tsv("outputs/calib.tsv") %>% 
  filter(b0 == 0.02, delta == 0.2) %>% 
  ggplot(aes(yday(date), simulated, col = a0, fill = a0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = as.factor(a0)), method = "lm", se = FALSE, 
              formula = y ~ poly(x, 20, raw = TRUE)) +
  theme_bw() +
  facet_wrap(~ site, scales = "free", nrow = 1) +
  scale_color_viridis_c("Year") +
  xlab("") + ylab("Simulated litterfall (MgC/ha/yr)") +
  scale_x_continuous(breaks = yday(as_date(paste0("2001-", 1:12, "-1"))),
                   labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  scale_color_viridis_c("a0 | b0 = 0.02, delta = 0.2") +
  scale_fill_viridis_c("a0 | b0 = 0.02, delta = 0.2") +
  theme(legend.position = "bottom")
ga1
```

```{r a0indices}
#| message: false
#| warning: false
#| fig-cap: "Effect of the leaf resistance parameter a on the relation between Litterfall peak date and ratioi across years at Paracou and Tapajos."
ga2 <- sim_metrics %>% 
  filter(b0 == 0.02, delta == 0.2, type == "simulated") %>% 
  ggplot(aes(peak_day, ratio, col = a0)) +
  geom_point() +
  theme_bw() +
  facet_wrap(~ site) +
  ylab("Simulated litterfall ratio") +
  xlab("Simulated litterfall peak day") +
  scale_color_viridis_c() +
  scale_color_viridis_c("a0 | b0 = 0.02, delta = 0.2") +
  theme(legend.position = "bottom")
ga2
```

The the tree height susceptibility parameter drives mainly the peak date but also the peak intensity for values above 0.1. Indeed low values close to 0.01 show a peak starting in September, whereas high values have a peak starting in December.

```{r b0}
#| message: false
#| warning: false
#| fig-cap: "Effect of the tree height susceptibility parameter b on the annual cycle from fortnightly litterfall for Paracou and Tapajos."
gb1 <- read_tsv("outputs/calib.tsv") %>% 
  filter(a0 == 0.2, delta == 0.2) %>% 
  ggplot(aes(yday(date), simulated, col = b0, fill = b0)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = as.factor(b0)), method = "lm", se = FALSE, 
              formula = y ~ poly(x, 20, raw = TRUE)) +
  theme_bw() +
  facet_wrap(~ site, scales = "free", nrow = 1) +
  scale_color_viridis_c("Year") +
  xlab("") + ylab("Simulated litterfall (MgC/ha/yr)") +
  scale_x_continuous(breaks = yday(as_date(paste0("2001-", 1:12, "-1"))),
                   labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  scale_color_viridis_c("b0 | a0 = 0.2, delta = 0.2") +
  scale_fill_viridis_c("b0 | a0 = 0.2, delta = 0.2") +
  theme(legend.position = "bottom")
gb1
```

```{r b0indices}
#| message: false
#| warning: false
#| fig-cap: "Effect of the tree height susceptibility parameter b on the relation between Litterfall peak date and ratioi across years at Paracou and Tapajos."
gb2 <- sim_metrics %>% 
  filter(a0 == 0.2, delta == 0.2, type == "simulated") %>% 
  ggplot(aes(peak_day, ratio, col = b0)) +
  geom_point() +
  theme_bw() +
  facet_wrap(~ site) +
  ylab("Simulated litterfall ratio") +
  xlab("Simulated litterfall peak day") +
  scale_color_viridis_c() +
  scale_color_viridis_c("b0 | a0 = 0.2, delta = 0.2") +
  theme(legend.position = "bottom")
gb2
```

The base variation in residence time multiplying factor for drought shedding seems to have a lighter influence on litterfall peak intensity and date, at least for default values of the two other parameters.

```{r delta}
#| message: false
#| warning: false
#| fig-cap: "Effect of the base variation in residence time multiplying factor for drought shedding delta on the annual cycle from fortnightly litterfall for Paracou and Tapajos."
gd1 <- read_tsv("outputs/calib.tsv") %>% 
  filter(b0 == 0.02, a0 == 0.2) %>% 
  ggplot(aes(yday(date), simulated, col = delta, fill = delta)) +
  geom_point(alpha = 0.5) +
  geom_smooth(aes(group = as.factor(delta)), method = "lm", se = FALSE, 
              formula = y ~ poly(x, 20, raw = TRUE)) +
  theme_bw() +
  facet_wrap(~ site, scales = "free", nrow = 1) +
  scale_color_viridis_c("Year") +
  xlab("") + ylab("Simulated litterfall (MgC/ha/yr)") +
  scale_x_continuous(breaks = yday(as_date(paste0("2001-", 1:12, "-1"))),
                   labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  scale_color_viridis_c("delta | a0 = 0.2, b0 = 0.02") +
  scale_fill_viridis_c("delta | a0 = 0.2, b0 = 0.02") +
  theme(legend.position = "bottom")
gd1
```

```{r deltaindices}
#| message: false
#| warngin: false
#| fig-cap: "Effect of the base variation in residence time multiplying factor for drought shedding delta on the relation between Litterfall peak date and ratioi across years at Paracou and Tapajos."
gd2 <- sim_metrics %>% 
  filter(b0 == 0.02, a0 == 0.2, type == "simulated") %>% 
  ggplot(aes(peak_day, ratio, col = delta)) +
  geom_point() +
  theme_bw() +
  facet_wrap(~ site) +
  ylab("Simulated litterfall ratio") +
  xlab("Simulated litterfall peak day") +
  scale_color_viridis_c() +
  scale_color_viridis_c("delta | a0 = 0.2, b0 = 0.02") +
  theme(legend.position = "bottom")
gd2
```

```{r fs02}
#| eval: false
g <- cowplot::plot_grid(ga1 + 
                          ggtitle("a0 | b0 = 0.02, delta = 0.2") +
                          theme(legend.position = "none"), 
                        gb1+ 
                          ggtitle("b0 | a0 = 0.2, delta = 0.2") +
                          theme(legend.position = "none"), 
                        gd1+ 
                          ggtitle("delta | a0 = 0.2, b0 = 0.02") +
                          theme(legend.position = "none"), 
                        ga2 +
                          scale_color_viridis_c(""),
                        gb2 +
                          scale_color_viridis_c(""), 
                        gd2 +
                          scale_color_viridis_c(""),
                        nrow = 2)
ggsave("figures/fs02.png", g, dpi = 300, width = 16, height = 8, bg = "white")
```

## Sensitivity to parameters

```{r pcc}
#| message: false
#| warning: false
t <- sim_metrics %>% filter(type == "simulated")
sens <- lapply(c("peak_day", "ratio"), function(var) 
  sensitivity::pcc(t[c("a0", "b0", "delta")], t[var], nboot = 100)$PCC %>%
                      rownames_to_column(var = "parameter") %>% 
    mutate(variable = var)) %>% 
  bind_rows()
```

```{r pcc_fig}
#| message: false
#| warning: false
g2 <- sens %>% 
  mutate(parameter = recode(parameter,
                            "a0" = "a[0]",
                            "b0" = "b[0]",
                            "delta" = "delta[0]")) %>% 
   mutate(variable = recode(variable,
                            "peak_day" = "Peak day",
                            "ratio" = "Ratio")) %>% 
  ggplot(aes(parameter, original, col = variable)) + 
  geom_point(position = position_dodge(0.4)) + 
  theme_bw() +
  coord_flip() + 
  geom_linerange(aes(ymin = `min. c.i.`, ymax = `max. c.i.`), 
                 position = position_dodge(0.4)) + 
  ylab("Partial Correlation Coefficients (PCC)") + xlab("") +
  theme(legend.position = "bottom") +
  scale_x_discrete(labels = scales::label_parse()) +
  scale_colour_discrete("")
g2
```

```{r}
#| eval: false
t <- data.frame(file = list.files("results/calib_pheno/")) %>% 
  rowwise() %>% 
  mutate(structure = list(read_tsv(file.path("results", "calib_pheno", file),
                                   col_select = c(3,5,7), skip = 4015, col_names = F) %>% 
                            tail(1))) %>% 
  unnest(structure) %>% 
  rename(abundance = X3, ba = X5, agb = X7) %>% 
  separate(file, c("site", "a0", "b0", "delta"), sep = "_", convert = TRUE)
sens <- lapply(c("abundance", "ba", "agb"), function(var) 
  sensitivity::pcc(t[c("a0", "b0", "delta")], t[var], nboot = 100)$PCC %>%
                      rownames_to_column(var = "parameter") %>% 
    mutate(variable = var)) %>% 
  bind_rows()
sens %>% 
  mutate(parameter = recode(parameter,
                            "a0" = "a[0]",
                            "b0" = "b[0]",
                            "delta" = "delta[0]")) %>% 
  ggplot(aes(parameter, original, col = variable)) + 
  geom_point(position = position_dodge(0.4)) + 
  theme_bw() +
  coord_flip() + 
  geom_linerange(aes(ymin = `min. c.i.`, ymax = `max. c.i.`), 
                 position = position_dodge(0.4)) + 
  ylab("Partial Correlation Coefficients (PCC)") + xlab("") +
  theme(legend.position = "bottom") +
  scale_x_discrete(labels = scales::label_parse()) +
  scale_colour_discrete("")
```

## Best parameters

We found similar $a=0.2$ and $b=0.015$ to be the best in Paracou and Tapajos, but slightly different $delta$ with $delta=0.1$ in Paracou and $delta=0.2$ in Tapajos. Except for delta in Tapajos all values are outside the limit of parameters explorations. The resulting litterfall dynamics of the simulations follows the intrannual pattern of observed litterfall with an early dry season peak (\~September Paracou, \~August Tapajos) of around 7 times the basal flux observed in the wet season. However, the model still have issues to simulated the intensity of the basal flux, especially in Tapajos.

```{r best}
#| message: false
#| warning: false
best <- sim_metrics %>% 
  ungroup() %>% 
  mutate(ratio = as.vector(scale(ratio)), 
         peak_day = as.vector(scale(peak_day))) %>% 
  gather(metric, value, -site, -a0, -b0, -delta, -year, -type) %>% 
  filter(metric %in% c("ratio", "peak_day")) %>% 
  mutate(metric = paste0(metric, "_", type)) %>% 
  select(-type) %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  na.omit() %>% 
  mutate(error = sqrt((ratio_observed - ratio_simulated)^2
                      +(peak_day_observed - peak_day_simulated)^2)) %>% 
  group_by(site, a0, b0, delta) %>% 
  summarise(RMSEP = sqrt(mean(error^2, na.rm = T))) %>% # should be normed
  group_by(site) %>% 
  slice_min(RMSEP) %>% 
  mutate(sim = paste0(site, "_", a0, "_", b0, "_", delta))
best %>% 
  knitr::kable(caption = "Best phenology parameters values according to calibration with litterfall data.")
```

```{r bestwithin}
#| message: false
#| warning: false
#| fig-cap: "Annual cycle from fortnightly litterfall for Paracou and Tapajos. Each thin line represents one year, the thick lines mean polynomial smoothing among years, the point field sampling dates, and rectangles in the background correspond to the site's climatological dry season."
g <- vroom::vroom("outputs/calib.tsv") %>% 
  mutate(sim = paste0(site, "_", a0, "_", b0, "_", delta)) %>% 
  filter(sim %in% best$sim) %>%
  select(-sim, -a0, -b0, -delta, -dt) %>% 
  gather(type, value, -site, -date) %>%  
  mutate(day = yday(date)) %>% 
  ggplot(aes(day, value)) +
  geom_rect(aes(xmin = day, xmax = end), fill  = "#fff4e0",
           ymin = -Inf, ymax = Inf, alpha = 0.5,
           data = data.frame(day = yday(as_date(c("2000-8-1", "2000-06-15"))),
                             value = -Inf,
                             end = yday(as_date(c("2000-12-1", "2000-11-1"))),
                             site = c("Paracou", "Tapajos"))) +
  geom_smooth(aes(col = type),
              method = "lm", formula = y ~ poly(x, 20, raw = TRUE), alpha = 0.5, se = FALSE) +
  geom_line(aes(col = type, group = paste0(year(date), type)), alpha = 0.6) +
  geom_point(aes(col = type), alpha = 0.7) +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_wrap(~ site, nrow = 2) +
  xlab("") + ylim(0, NA) +
  scale_x_continuous(breaks = yday(as_date(paste0("2001-", 1:12, "-1"))),
                   labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  scale_color_manual("", values = as.vector(cols[c("obs", "sim")]), 
                     labels = c("Litter traps", "TROLL")) +
  ylab(expression(Litterfall~"["~MgC~ha^{-~1}~yr^{-~1}~"]")) 
ggsave("figures/f06.png", g, dpi = 300, width = 5, height = 5, bg = "white")
g
```

```{r}
vroom::vroom("outputs/calib.tsv") %>% 
  mutate(sim = paste0(site, "_", a0, "_", b0, "_", delta)) %>% 
  filter(sim %in% best$sim) %>%
  left_join(vroom::vroom("outputs/calib_test_ll_isa.tsv") %>% 
  rename(simulated_ll = simulated)) %>% 
  select(-sim, -a0, -b0, -delta, -dt) %>% 
  gather(type, value, -site, -date) %>%  
  mutate(day = yday(date)) %>% 
  ggplot(aes(day, value)) +
  geom_rect(aes(xmin = day, xmax = end), fill  = "#fff4e0",
           ymin = -Inf, ymax = Inf, alpha = 0.5,
           data = data.frame(day = yday(as_date(c("2000-8-1", "2000-06-15"))),
                             value = -Inf,
                             end = yday(as_date(c("2000-12-1", "2000-11-1"))),
                             site = c("Paracou", "Tapajos"))) +
  geom_smooth(aes(col = type),
              method = "lm", formula = y ~ poly(x, 20, raw = TRUE), alpha = 0.5, se = FALSE) +
  geom_line(aes(col = type, group = paste0(year(date), type)), alpha = 0.6) +
  geom_point(aes(col = type), alpha = 0.7) +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_wrap(~ site, nrow = 2) +
  xlab("") + ylim(0, NA) +
  scale_x_continuous(breaks = yday(as_date(paste0("2001-", 1:12, "-1"))),
                   labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D"))
  ylab(expression(Litterfall~"["~MgC~ha^{-~1}~yr^{-~1}~"]")) 
```

```{r}
vroom::vroom("outputs/calib.tsv") %>% 
  mutate(sim = paste0(site, "_", a0, "_", b0, "_", delta)) %>% 
  filter(sim %in% best$sim) %>%
  select(-sim, -a0, -b0, -delta, -dt) %>% 
  mutate(simulated = simulated*2) %>% 
  gather(type, value, -site, -date) %>%  
  mutate(day = yday(date)) %>% 
  ggplot(aes(day, value)) +
  geom_rect(aes(xmin = day, xmax = end), fill  = "#fff4e0",
           ymin = -Inf, ymax = Inf, alpha = 0.5,
           data = data.frame(day = yday(as_date(c("2000-8-1", "2000-06-15"))),
                             value = -Inf,
                             end = yday(as_date(c("2000-12-1", "2000-11-1"))),
                             site = c("Paracou", "Tapajos"))) +
  geom_smooth(aes(col = type),
              method = "lm", formula = y ~ poly(x, 20, raw = TRUE), alpha = 0.5, se = FALSE) +
  geom_line(aes(col = type, group = paste0(year(date), type)), alpha = 0.6) +
  geom_point(aes(col = type), alpha = 0.7) +
  theme_bw() +
  theme(legend.position = "bottom") +
  facet_wrap(~ site, nrow = 2) +
  xlab("") + ylim(0, NA) +
  scale_x_continuous(breaks = yday(as_date(paste0("2001-", 1:12, "-1"))),
                   labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
  scale_color_manual("", values = as.vector(cols[c("obs", "sim")]), 
                     labels = c("Litter traps", "TROLL")) +
  ylab(expression(Litterfall~"["~MgC~ha^{-~1}~yr^{-~1}~"]")) 
```

```{r bestamng}
#| message: false
#| fig-cap: "Fortnightly litterfall for Paracou and Tapajos simulated in TROLL and measured with litter traps."
vroom::vroom("outputs/calib.tsv") %>% 
  mutate(sim = paste0(site, "_", a0, "_", b0, "_", delta)) %>% 
  filter(sim %in% best$sim) %>%
  select(-sim, -a0, -b0, -delta, -dt) %>% 
  gather(type, value, -site, -date) %>% 
  ggplot(aes(date, value, col = type)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~ site, scales = "free_x") +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_discrete("", labels = c("field", "TROLL")) +
  xlab("") + 
  scale_color_manual("", values = as.vector(cols[c("obs", "sim")]), 
                     labels = c("Litter traps", "TROLL")) +
  ylab(expression(Litterfall~"["~MgC~ha^{-1}~yr^{-1}~"]")) 
```
