```{r setup}
#| include: false 
rm(list = ls())
library(tidyverse) 
library(sf) 
library(leaflet)
```

# Introduction {.unnumbered}

The aim of these analyses is to evaluate the TROLL version 4 simulations of daily fluxes and the final structure and composition of the forest in relation to field observations and remote sensing at two Amazonian sites.

## Study sites

We have four sites with long-term eddy flux tower in Amazonia but we were only able to collect data from 2 in Brazil and 1 in French Guiana with free access. As BR-Sa3 also known as Km83 is in a logged forest on a shorter period and with lower rainfall registered in the flux tower we will focus on GF-Guy, hereafter Paracou, and BR-Sa1, hereafter Tapajos.

```{r sites}
tibble(
  site = c("Paracou", "Tapajos"),
  latitude = c(5.27877, -2.85667),
  longitude = c(-52.92486, -54.9588900),
) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(label = ~site, labelOptions = labelOptions(noHide = T))
```

## Poeple

-   Sylvain Schmitt (sylvain.schmitt\@cirad.fr)
-   Isabelle Maréchaux (isabelle.marechaux\@inrae.fr)
-   Jérôme Chave (jerome.chave\@univ-tlse3.fr)
-   Fabian J. Fischer (f.fischer\@bristol.ac.uk)
-   Damien Bonal (damien.bonal\@inrae.fr)
-   Benoit Burban(benoit.burban\@inrae.fr)
-   Nicolas Barbier (nicolas.barbier\@ird.fr)
-   Gregoire Vincent (gregoire.vincent\@ird.fr)
-   Philippe Verley (philippe.verley\@ird.fr)
-   *James Ball* *(jgcb3\@cam.ac.uk)?*
-   Géraldine Derroire (geraldine.derroire\@cirad.fr)
-   Giacomo Selan (giacomo.sellan\@cirad.fr)
-   *Jérémy Lischtein ()?*
-   Daniela Nemetschek (Daniela.Krebber\@ecofog.gf)
-   Marion Boisseaux (marionboisseaux\@gmail.com)
-   Camille Ziegler (camille.ziegler9\@gmail.com)
