```{r setup}
#| include: false 
rm(list = ls())
library(tidyverse) 
library(sf) 
library(leaflet)
```

# Introduction {.unnumbered}

The aim of these analyses is to evaluate the TROLL 4.0 simulations of daily fluxes and the final structure and composition of the forest in relation to field observations and remote sensing at two Amazonian sites.

## Study sites

We have four sites with long-term eddy flux tower in Amazonia but we were only able to collect data from 2 in Brazil and 1 in French Guiana with free access. As BR-Sa3 also known as Km83 is in a logged forest on a shorter period and with lower rainfall registered in the flux tower we will focus on GF-Guy, hereafter Paracou, and BR-Sa1, hereafter Tapajos.

```{r sites}
tibble(
  site = c("Paracou", "Tapajos"),
  latitude = c(5.27877, -2.85667),
  longitude = c(-52.92486, -54.9588900),
) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(label = ~site, labelOptions = labelOptions(noHide = T))
```

## Poeple

-   Sylvain Schmitt
-   Fabian J. Fischer
-   James G. C. Ball
-   Nicolas Barbier
-   Marion Boisseaux
-   Damien Bonal
-   Benoit Burban
-   Xiuzhi Chen
-   Géraldine Derroire
-   Jeremy W. Lichstein
-   Daniela Nemetschek
-   Natalia Restrepo-Coupe
-   Scott Saleska
-   Giacomo Sellan
-   Philippe Verley
-   Gregoire Vincent
-   Camille Ziegler
-   Jérôme Chave
-   Isabelle Maréchaux
