---
title: "Species data"
author: Sylvain Schmitt -
date: today
date-format: medium
format: 
  gfm:
    toc: true
    toc-depth: 2
---

This folder contains species allometric parameters data (mainly height and dbh) and functional traits. Data were gathered from previous studies including:

-   @jucker2022
-   @maréchaux2015
-   @guillemot2022
-   @vleminckx2021

TROLL version 4 species data were gathered from:

-   TROLL version 3 species (partly from @baraloto2010, described in @maréchaux2017, and available in \@schmitt2023)
-   @maréchaux2015
-   @maréchaux2019
-   @nemetschek2024
-   @schmitt2023a
-   Boisseaux *et al.* submitted
-   Ziegler *et al.* (which?)

Tapajos functional traits data were also retrieved by Jérôme Chave (pers. com.) from the TRY database [@Kattge2019].

```{r get}
#| eval: false
library(tidyverse)
write_tsv(rcontroll::TROLLv4_species, "troll.tsv")
download.file("https://zenodo.org/records/6637599/files/Tallo.csv?download=1", "tallo.csv")
download.file("https://besjournals.onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1111%2F1365-2435.12452&file=fec12452-sup-0002-AppendixS1.xls", "marechaux.xls")
download.file("https://dataverse.cirad.fr/dataset.xhtml?persistentId=doi:10.18167/DVN1/HLBVJG&version=1.1#", "guillemot.xls")
download.file("https://nsojournals.onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1111%2Foik.08284&file=oik13191-sup-0002-AppendixS2.xlsx", "vleminckx.xlsx")
# TRY extract by Jérôme Chave
```

```{r, comment=""}
fs::dir_tree()
```
